Bit Complexity for Computation of one Point in each Connected Component of a Smooth Real Hypersurface

### 2020-01-31 main.tex
Improved code readability

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
We provide bit complexity estimates for the computation of at least one point in each connected component of a smooth real hypersurface. This problem is a basic and important subroutine used in semi-algebraic geometry. For instance, it is used in determining an upper bound on the number of connected components of a real hypersurface. The algorithm requires some genericity properties, which we ensure through random changes of variables that produce generic coordinate systems. In order to ensure these properties with arbitrarily high probability, we developed a new quantitative extension of Thom's weak transversality theorem, and a quantitative Noether normalization statement for polar varieties. These statements enable us to give a precise probability analysis of a modular algorithm, where the algorithm is monte carlo and we can guarantee correctness with arbitrarily high probability. 
\par
Algorithms for deciding connectivity queries in real algebraic geometry have been developed in a series of papers \cite{a,b,c,d}. However, these algorithms are based in the arithmetic complexity model, and bit complexity questions have been left for future work.  Our results therefore begin to fill a missing gap in the research. 
%Furthermore, as bit complexity provides a far more realistic model of computation compared to arithmetic complexity, which only counts each operation at unit cost whereas bit complexity provide more information for both time and space resources that are needed in practice, it is important that this missing gap is filled. 
%As bit complexity provides a far more realistic model of computation compared to arithmetic complexity, which only counts each operation at unit cost, our contribution is necessary and important. Indeed, bit estimates provide more information for both time and space resources that are needed in practice. 
\end{abstract}

@@

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Real algebraic geometry, weak transversality, Noether position, bit complexity}
%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.

@@

%%%%%% Introduction

@@

%%%%%% Problem statement

@@

%%%%%% Motivation

@@

%Indeed, bit estimates provide more information for both time and space resources that are needed in practice.  

@@

%%%%%% Main result

@@

%%%%%% Notation and Preliminaries

@@

%%%%%% Algebraic Sets

@@

%%%%%% Bit-size, height and degree

@@

%%%%%% Data structures

@@

%%%%%% Regular and singular points

@@

%%%%%%%%%%%%%%%% Critical points of polynomial mappings

@@

%%%%%% The zariski-tangent space

@@

%and $\jac(\textit{\textbf{f}},i)$ the truncated Jacobian matrix
%\[ 
%\bbm 
%\frac{\partial f_1}{\partial X_{i+1}} \hdots \frac{\partial f_1}{\partial X_n} \\
%\frac{\partial f_N}{\partial X_{i+1}} \hdots \frac{\partial f_N}{\partial X_n}
%\ebm,~\textrm{for }i \in \{1,\hdots,n-1\}.\]

@@

%%%%%% Polar Varieties

@@

%%%%%% Changes of Variables

@@

%%%%%% Main Algorithm 

@@

\nl Use \cite{4} to compute the partial derivatives $\frac{\pa f^A}{\pr X_{i+1}},\hdots,\frac{\pa f^A}{\pr X_n}$\;

@@

%%%%%% Genericity Properties

@@

Consider $f \in \mathbb{Q}[X_1,\hdots,X_n]$ squarefree with total degree $D$ and $V(f)$ smooth. We say that $V$ is in \textit{Noether position} for $\pi_D$ when the extension \[\C[X_1,\hdots,X_D] \rightarrow \C[X_1,\hdots,X_n]/I(V)\] is injective and integral.  Then, for any $\xb \in \C^D,$ the fiber 

@@

%%%%%% A Quantitative Extension of Thom's Weak Transversality

@@

\[\deg \Delta \leq \deg \Delta^{'} \leq d^n.

@@

\begin{align*}M(\xb,\vt)&:= \jac_{(\xb,\vt)}(\Phi;X)\\

@@

\ebm.\end{align*}

@@

 \[J := \jac_{(\xb,\vt)}(\Phi)[-;1,n].\]

@@

\[\overline{\pi_{n+\dt}~(\overline{Z^o})} = \overline{\pi_{n+\dt}~(Z^o)}.\] 

@@

\[\pi_{n+\dt}~(Z^{o}) \subset \D^{'}~(\textrm{why?}).\] 

@@

\[\overline{\pi_{n+\dt}~(Z^o)} \subset \overline{\pi_{n+\dt}~(\overline{Z^o})}.

@@

\[\overline{\pi_{n+\dt}~(\overline{Z^o})} = V(\mathscr{I}),

@@

\[(f(Ax),(\grad(f)\cdot 

@@

\[(f,\grad(f)\cdot 

@@

\end{align*} where 

@@

\end{proof}
Since $\deg \Phi \leq D,$ it follows by Theorem 4.2 that a hypersurface $\Delta \subset \C^{n^2}$ exists, of degree at most $D^n,$ with the property that, if $A \in \C^{n^2}-\Delta,$ then $0$ is a regular value of $\Phi_{A}$, which means that 
\[
\jac_\textit{\textbf{x}} \Phi_A 
= \jac_\xb \left(f^A,\frac{\pa f^A}{\pa X_{i+1}},\hdots,\frac{\pa f^A}{\pa X_n}\right)
\] has full rank for all $\xb \in V(f^A)$ ($\cap \R^n$?). Therefore, by the Jacobian criterion \cite[Theorem 16.19]{ECA}, $f^A$ satisfies $\textbf{H}_i^{'}(1),\textbf{H}_i^{'}(2)$ and $\textbf{H}_i^{'}(3)$, for each $i \in \{1,\hdots,n\}.$
%
%
%
%
%
%%%%%%% Proof of $\textbf{H}_i^{'}(1),\textbf{H}_i^{'}(2)$ and $\textbf{H}_i^{'}(3)$
\section{Proof of $\textbf{H}_i^{'}(4),\textbf{H}_i^{'}(5)$ and $\textbf{H}_i^{'}(6)$}
%
\begin{theorem}
There exists a Zariski closed set $\Delta_2 \subset \C^{n^2}$ of degree at most $D^{\textrm{?}}$ with the property that, if $A \in \C^{n^2} - \Delta_2$ then for each $i \in \{1,\hdots,n\},$ $f^A$ satisfies $\textbf{H}_i^{'}(4),\textbf{H}_i^{'}(5)$ and $\textbf{H}_i^{'}(6)$.
\end{theorem}
%
\subsection{Sketch of the proof of Theorem 7.1}
%
Let 
\[
Y_i = X_i^{A} = [b_i] \cdot \bbm X_1\\ \vdots \\ X_n \ebm, 
\]
where $b_i$ is the i-th row of $A^{-1}$. For $i \in \{1,\hdots,n\}$, we want to show that the Jacobian of the system of polynomials
\[
Y_1(\xb),\hdots,Y_{i-1}(\xb), f(\xb),  \frac{\pa f(A\xb)}{\pa X_{i+1}},\hdots,\frac{\pa f(A\xb)}{\pa X_n}
\]
has full rank, at all points $\xb \in \C^n$ with 
\[
X_1(\xb)=\hdots,X_{i-1}(\xb)=f(\xb)= \frac{\pa f(xb)}{\pa X_{i+1}}=\hdots = \frac{\pa f(\xb)}{\pa X_{n}}=0.
\]
Then we could use Theorem 5.1 to obtain a hypersurface with the desired properties. Let $\Phi$ be the mapping 
\begin{align*}
    \Phi: \C^n \times \C^{n^2} &\rightarrow \C^{n+1} \\
           (\xb,A) &\mapsto 
       (Y_1,\hdots,Y_i,f(\xb),\grad(f(\xb))\cdot A_{i+1},\hdots \\
       \hdots&~,\grad(f(\xb))\cdot A_{n}\right)),
\end{align*}
We cannot directly define the appropriate polynomial mapping, as $Y_j(A\xb)$ is rational in $A^{-1}.$ To circumvent this issue, we can alternativly define a polynomial mapping $\Psi$ as  
\begin{align*}
    \Psi: \C^n \times \C^{n^2} \times \C &\rightarrow \C^{n+2} \\
           (\xb,A,T) &\mapsto 
       (\Yt_1,\hdots,\Yt_i,f(\xb),\grad(f(\xb))\cdot A_{i+1},\hdots \\
\hdots,\grad(f(\xb))\cdot A_{n}\right),&~T\cdot \det (A) - 1),
\end{align*}
where $\Yt_j=TY_j$ with denominators cleared. Now, look at the points in $\Psi^{-1}(0).$ Here, $\det A \not = 0$ and therefore invertible, and therefore at all points in $\Psi^{-1}(0)$ the Jacobian matrix has full rank? 
\begin{proposition}
For all $(\xb,A) \in \Psi^{-1}(0),$ the Jacobian matrix of $\Psi$ at $(\xb,A)$ has full rank (i.e. $\Psi$ is transverse to $\{0\}).$
\end{proposition}
\begin{proof}

@@

@@

%%%%%% Quantitative Noether Position

@@

%%%%% Theorem

@@

%%%%% Proposition

@@

%%%%%% Proof of Theorem 3.1

@@

%%%%%% Bit-complexity

@@

%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-base}
\end{document}
\endinput
%% End of file `sample-xelatex.tex'.

>>>>

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
We provide bit complexity estimates for the computation of at least one point in each connected component of a smooth real hypersurface. This problem is a basic and important subroutine used in semi-algebraic geometry. For instance, it is used in determining an upper bound on the number of connected components of a real hypersurface. The algorithm requires some genericity properties, which we ensure through random changes of variables that produce generic coordinate systems. In order to ensure these properties with arbitrarily high probability, we developed a new quantitative extension of Thom's weak transversality theorem, and a new quantitative Noether normalization statement for polar varieties. These statements enable us to give a precise probability analysis of a monte carlo algorithm, where we can guarantee correctness with arbitrarily high probability. 
\par
Algorithms for deciding connectivity queries in real algebraic geometry have been developed in a series of papers \cite{a,b,c,d}. However, these algorithms are based in the arithmetic complexity model, and bit complexity questions were left for future work.  Our results therefore begin to fill a missing gap in the research. 
%Furthermore, as bit complexity provides a far more realistic model of computation compared to arithmetic complexity, which only counts each operation at unit cost whereas bit complexity provide more information for both time and space resources that are needed in practice, it is important that this missing gap is filled. 
%As bit complexity provides a far more realistic model of computation compared to arithmetic complexity, which only counts each operation at unit cost, our contribution is necessary and important. Indeed, bit estimates provide more information for both time and space resources that are needed in practice. 
\end{abstract}

@@

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Real algebraic geometry, transversality, Noether normalization, bit complexity}
%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

\nl Use \cite{AD} to compute the partial derivatives $\frac{\pa f^A}{\pr X_{i+1}},\hdots,\frac{\pa f^A}{\pr X_n}$\;

@@

@@

Consider $f \in \mathbb{Q}[X_1,\hdots,X_n]$ squarefree with total degree $D$ and $V(f)$ smooth. We say that $V$ is in \textit{Noether position} for $\pi_D$ when the extension 
\C[X_1,\hdots,X_D] \rightarrow \C[X_1,\hdots,X_n]/I(V)
is injective and integral.  Then, for any $\xb \in \C^D,$ the fiber 

@@

@@

\deg \Delta \leq \deg \Delta^{'} \leq d^n.

@@

M(\xb,\vt)&:= \jac_{(\xb,\vt)}(\Phi;X)\\

@@

@@

 J := \jac_{(\xb,\vt)}(\Phi)[-;1,n].

@@

\overline{\pi_{n+\dt}~(\overline{Z^o})} = \overline{\pi_{n+\dt}~(Z^o)}.

@@

\pi_{n+\dt}~(Z^{o}) \subset \D^{'}~(\textrm{why?}).

@@

\overline{\pi_{n+\dt}~(Z^o)} \subset \overline{\pi_{n+\dt}~(\overline{Z^o})}.

@@

\overline{\pi_{n+\dt}~(\overline{Z^o})} = V(\mathscr{I}),

@@

(f(Ax),(\grad(f)\cdot 

@@

(f,\grad(f)\cdot 

@@

where 

@@

@@

Since $\deg \Phi \leq D,$ it follows by Theorem 4.2 that a hypersurface $\Delta \subset \C^{n^2}$ exists, of degree at most $D^n,$ with the property that, if $A \in \C^{n^2}-\Delta,$ then $0$ is a regular value of $\Phi_{A}$, which means that 
\jac_\textit{\textbf{x}} \Phi_A 
= \jac_\xb \left(f^A,\frac{\pa f^A}{\pa X_{i+1}},\hdots,\frac{\pa f^A}{\pa X_n}\right)
has full rank for all $\xb \in V(f^A)$ ($\cap \R^n$?). Therefore, by the Jacobian criterion \cite[Theorem 16.19]{ECA}, $f^A$ satisfies $\textbf{H}_i^{'}(1),\textbf{H}_i^{'}(2)$ and $\textbf{H}_i^{'}(3)$, for each $i \in \{1,\hdots,n\}.$
\section{Proof of $\textbf{H}_i^{'}(4),\textbf{H}_i^{'}(5)$ and $\textbf{H}_i^{'}(6)$}
There exists a Zariski closed set $\Delta_2 \subset \C^{n^2}$ of degree at most $D^{\textrm{?}}$ with the property that, if $A \in \C^{n^2} - \Delta_2$ then for each $i \in \{1,\hdots,n\},$ $f^A$ satisfies $\textbf{H}_i^{'}(4),\textbf{H}_i^{'}(5)$ and $\textbf{H}_i^{'}(6)$.
\subsection{Sketch of the proof of Theorem 7.1}
Let 
Y_i = X_i^{A} = [b_i] \cdot \bbm X_1\\ \vdots \\ X_n \ebm, 
where $b_i$ is the i-th row of $A^{-1}$. For $i \in \{1,\hdots,n\}$, we want to show that the Jacobian of the system of polynomials
Y_1(\xb),\hdots,Y_{i-1}(\xb), f(\xb),  \frac{\pa f(A\xb)}{\pa X_{i+1}},\hdots,\frac{\pa f(A\xb)}{\pa X_n}
has full rank, at all points $\xb \in \C^n$ with 
X_1(\xb)=\hdots,X_{i-1}(\xb)=f(\xb)= \frac{\pa f(xb)}{\pa X_{i+1}}=\hdots = \frac{\pa f(\xb)}{\pa X_{n}}=0.
Then we could use Theorem 5.1 to obtain a hypersurface with the desired properties. Let $\Phi$ be the mapping 
    \Phi: \C^n \times \C^{n^2} &\rightarrow \C^{n+1} \\
       (Y_1,\hdots,Y_i,f(\xb),\grad(f(\xb))\cdot A_{i+1},\hdots \\
       \hdots&~,\grad(f(\xb))\cdot A_{n})),
We cannot directly define the appropriate polynomial mapping, as $Y_j(A\xb)$ is rational in $A^{-1}.$ To circumvent this issue, we can alternativly define a polynomial mapping $\Psi$ as  
    \Psi: \C^n \times \C^{n^2} \times \C &\rightarrow \C^{n+2} \\
           (\xb,A,T) &\mapsto 
       (\Yt_1,\hdots,\Yt_i,f(\xb),\grad(f(\xb))\cdot A_{i+1},\hdots \\
\hdots,\grad(f(\xb))\cdot A_{n}),&~T\cdot \det (A) - 1),
where $\Yt_j=TY_j$ with denominators cleared. Now, look at the points in $\Psi^{-1}(0).$ Here, $\det A \not = 0$ and therefore invertible, and therefore at all points in $\Psi^{-1}(0)$ the Jacobian matrix has full rank? 
For all $(\xb,A) \in \Psi^{-1}(0),$ the Jacobian matrix of $\Psi$ at $(\xb,A)$ has full rank (i.e. $\Psi$ is transverse to $\{0\}).$

@@

@@

@@

@@

@@

@@

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-base}
%%
\end{document}
\endinput
%%
%% End of file `sample-xelatex.tex'.
